% CHAPTER 1
\chapter{CONCLUSION}
\label{chp:7}
This thesis study highlights the problem of reduction in effective grid inertia that is caused by increasing renewable energy penetration. Since the renewable energy systems that are connected to grid with a power electronics interface decouples rotational speed and grid frequency, the inertia of these units do not contribute the frequency stability of the grid. However, it is shown that the reduction in the grid inertia can be compensated with inertial support implementation in the wind turbines with FSPC. It is even possible to improve the frequency stability of the grid with the implementation of synthetic inertia method that can emulate higher inertia constant than actual inertia constant existing in the turbinefeasibility . \par
\section{Final Remarks}
The inertial support implementations either frequency dependent or not can be provided with the utilization of the additional energy in the renewable energy systems. The increase in the active power is found to be limited by the converter's power rating meanwhile the available kinetic energy restricts the support duration. This thesis demonstrates the relation between the available increase in the active power and wind speed. It is observed that the increase in the active power is limited as 0.1pu in high wind speeds. However, the additional active power up to 0.45pu can be injected to grid when the wind speed is lower than 10m/s. The shortest support interval is slightly above 1 second and observed in the wind speed 3m/s. Therefore, it is concluded that the kinetic energy affects the support duration but the increase in active power is limited by wind speeds and converter capability. \par
For maximum RoCoF of 0.25Hz/s, the studied wind turbine can emulate the inertial constant up to 32seconds. Nonetheless, the inertia constant H=10s is applicable to whole wind speed range. By utilizing the synthetic inertia implementation H=10s, the effective kinetic energy existing in the grid inertia can be increased by 8\% in the Turkish electricity grid. Besides, the reduction in the effective grid inertia constant can be compensated with the synthetic inertia implementation of 10 seconds. Consequently, the share of the renewable energy can be increased without deteriorating the effective grid inertia.\par
The additional active power in the low and medium wind speed ranges are provided with the kinetic energy extraction from the turbine inertia. However, the additional power can be released from the turbine capacity by decreasing the pitch angle. In both cases, the active power can be increased within 20ms by considering the electrical time constant. However, the detection of the grid RoCoF and calculation of the required power and current references bring additional delays. Therefore, the real implementation on a wind turbine might be less effective compared to simulation case in which only the electrical time constant is under consideration.\par
This study also evaluates the economical perspective of the inertial support. The payment to inertia support provision is important to convince the energy provider for implementing the grid supporting applications. It is observed that the payment by considering the additional power is insignificant compared to daily average profit of the owner. In contrary, the significant amount of additional payment can be obtained as long as incentives are paid to renewable energy units for grid supporting applications. In this case, the payment is regardless from the additional energy but for the availability of the unit to the synthetic inertia implementation. In this way, the energy providers can be convinced to implement grid supporting applications that will increase the stability of the electricity grid.
\newpage
\section{Future Work}
The following issues can be further studied in detail:
\begin{itemize}
	\item Fast inertial support implementation has huge potential to increase the active power. However, as soon as the support period finishes, the reduction in the active power resembles a second load connection to system especially in the weak power systems. Therefore, the system is exposed to a secondary dip in the frequency. This is why the amount of additional power should be adjusted according to the grid frequency. Hence, an inertia support index can be developed to reshape the active power output of the wind turbine. 
	\item The increase percentage of fast inertial support is not a function of the grid parameters. The index to be constructed should also determine the increase percentage based on the pre-determined values. 
	\item The effects of the fast inertial support on the DC-link voltage might be better investigated with more realistic modelling. The effectiveness of the new modelling might be tested in the wind turbine emulators. 
	\item According to \cite{Altin2018}, increasing active power by 30\% according to the available power causes significant increases in  tower, yaw bearing and main bearing bending moments. Therefore, the effect of the inertial support in the turbine structure can be investigated in detail. 
	\item When the inertial support is required in the high wind speeds, support power might reach to the limit of the converter power rating. The extra heat and mechanical stresses might also be tested in wind turbine emulators. 
\end{itemize}